---
title: "Share of Search and the brand funnel"
subtitle: "Can search data complement or substitute survey-based brand tracking? A pilot study across 50 Australian categories"
author: "Zain Raza"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 4.8)
source(file.path("R", "functions.R"))
PAL <- c("#2a78d6", "#eb6834", "#1baf7a", "#eda100")
theme_set(theme_minimal(base_size = 12) +
          theme(panel.grid.minor = element_blank(),
                panel.grid.major = element_line(color = "grey92"),
                plot.title.position = "plot"))
```

```{r loading}
# A brand needs a minimum average Trends index to be usable. Trends reports
# 0-100 within each request, so a brand averaging under half a point can't
# support a share calculation. Section 8 shows a higher bar changes little.
VOL_MIN <- 0.5

survey <- load_survey()            # 50 categories of funnel metrics
trends <- load_trends("trends")    # Google Trends, collected in pull/

vol <- trends |>
  group_by(category_name, brand_name) |>
  summarise(mean_idx = mean(trends_index, na.rm = TRUE), .groups = "drop")
searchable <- filter(vol, mean_idx >= VOL_MIN)

# The analysis table: keep usable brands, turn search volume into share of
# search within each category-month, join the survey, do the same for the
# funnel metrics so everything is a share of its category.
panel <- trends |>
  semi_join(searchable, by = c("category_name", "brand_name")) |>
  group_by(category_name, month) |>
  mutate(sos = trends_index / sum(trends_index, na.rm = TRUE)) |>
  ungroup() |>
  inner_join(survey, by = c("category_name", "brand_name", "month" = "wave_date")) |>
  group_by(category_name, month) |>
  mutate(soc = consideration / sum(consideration, na.rm = TRUE),
         sop = preference / sum(preference, na.rm = TRUE)) |>
  ungroup()

# Only the 100-shuffle placebo is read from disk — it takes two minutes, so it
# runs in analysis/03_leadlag_placebo.R. Everything else computes here.
out <- function(f) readr::read_csv(file.path("out", f), show_col_types = FALSE)
```

# Summary {-}

Tracksuit's survey asks people what they *say*: have you heard of this brand,
would you consider it, do you prefer it. Search asks nothing — it records what
people *do*. This study asks how much of the saying the doing can replace.

**Three findings.**

**1. Search ranks brands like the survey does in some categories, not in
others.** So the product isn't a metric, it's a qualification check: which
categories can carry a search-based number before it reaches a client.

**2. Search can't replace the survey for month-to-month tracking — because of
the survey's sample size, not search's quality.** At ~200 respondents a month,
consideration carries a margin of error of about ±3 points. In the typical
category only a quarter of the monthly movement is real. Nothing correlates
well with a number that unsteady.

Scale is the real trade. Surveys scale, but every extra brand, category or
country costs sample, and precision comes from sample *per cell* — so scaling
out makes each number shakier, not steadier. Search scales the opposite way:
adding brands and countries is nearly free and it never runs out of
respondents. What it costs is control. You can't ask it a question and you
can't cut it by audience.

**3. The useful signal is the gap between the two, not the correlation between
them.** Brands searched for more than their consideration share implies gain
consideration, and preference, over the next six months. Brands searched for
less lose it.

**Recommendation.** Ship search as a directional early-warning layer on the
funnel. The survey says where a brand stands; the gap says which way it's
heading between waves. It won't replace the survey — no demographics, no funnel
stages, no absolute levels. And it can't be assembled casually: seven badly
matched brand names swung my headline result by 15 points. That's a data
engineering job, not a research one.

> The commercial case isn't cheaper tracking. It's faster tracking.

Clients wait a month for a number that mostly moves on noise. A search-gap flag
gives them a direction between waves, for free. I'd put it in front of clients
in the categories where it holds up and learn from what they do with it, rather
than perfecting every category first. A client who acts on a search signal and
sees the next wave confirm it teaches us more than another month of internal
validation.

# 1. The question, in funnel terms

The funnel is awareness → consideration → preference. Each step is a claim a
respondent makes about themselves.

A search is a behaviour, and it sits between the first two steps. You can't
search for a brand you've never heard of, so search needs awareness. Typing a
brand name takes more effort than recognising a logo, so it implies something
closer to intent. If that's right, share of search reads the top half of the
funnel behaviourally — and moves before the survey, because people act before
they're asked.

Two product questions, different bars:

| Question | If yes, search is | What has to be true |
|---|---|---|
| Does search rank brands like the funnel does? | a benchmarking layer | search ranks brands like consideration |
| Does search move *before* the funnel? | an early-warning layer | today's search predicts tomorrow's consideration |

Theory and assumptions are in [`assumptions.md`](assumptions.md).

# 2. The data

Google Trends for every brand in all 50 Australian categories, September 2021
to March 2025, matched to the survey waves. **Share of search** is a brand's
index divided by the total for its category, which puts it on the same footing
as share of consideration.

```{r data-summary}
tibble(categories = n_distinct(panel$category_name),
       brands = n_distinct(paste(panel$category_name, panel$brand_name)),
       months = n_distinct(panel$month),
       `brands with too little search` = nrow(vol) - nrow(searchable))
```

One number is a finding on its own. Of the 507 brands tracked here, 97 — about
1 in 5 — never register enough search volume to measure. Real brands, real
consideration scores, no search. That settles the substitution question on its
own: you can't track brands that are invisible.

Collecting this was most of the work and none of the insight. Google caps
comparisons at five terms per request, so 507 brands onto one scale needs
stitched requests plus heavy disambiguation — "Koala" is an animal, "Mecca" is
a city, "Ultimate" is a yoghurt. I used AI tooling throughout: the collection
pipeline, resolving names to Google entity IDs, reviewing the log. That was a
project in itself a few years ago; doing it in an afternoon is why this covers
50 categories instead of five. The run takes 90 minutes against rate limits, so
the raw data is committed — nobody needs to re-run it. Mechanics in Appendix A.

# 3. How steady is the survey itself?

Each brand's monthly score rests on the ~200 people who qualified for that
category. That carries a margin of error of about ±3 points, so a brand moving
25% to 28% is well inside sampling noise.

Measured per brand: take how much consideration actually moved, subtract what
sampling alone would produce, and what's left is real movement. One brand with
its sampling margin drawn around it:

```{r noise-example, fig.height=4}
eg <- survey |>
  filter(category_name == "Fast Food", brand_name == "KFC") |>
  arrange(wave_date) |>
  mutate(se = se_prop(consideration, base_n),
         q = zoo::rollmean(consideration, 3, fill = NA, align = "right"))

ggplot(eg, aes(wave_date, consideration)) +
  geom_ribbon(aes(ymin = consideration - 1.96 * se, ymax = consideration + 1.96 * se),
              fill = "grey86") +
  geom_line(color = "grey55", linewidth = .5) +
  geom_line(aes(y = q), color = PAL[1], linewidth = 1.1) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Consideration",
       title = "One brand's monthly consideration, with its margin of error",
       subtitle = "Grey: where the score could sit on sampling alone. Blue: the three-month average I use instead.")
```

This isn't a criticism of the survey — it's about resolution. Ask 200 people
and you can see a ten-point gap between brands; you can't see a two-point move
in one brand month to month. So I compare quarters, not months.

Across all brands, by category:

```{r noise, fig.height=4.4}
nd <- noise_decomposition(survey, "consideration")
nd_cat <- nd |> filter(waves >= 12) |> group_by(category_name) |>
  summarise(real = mean(reliability, na.rm = TRUE), .groups = "drop")

ggplot(nd_cat, aes(real)) +
  geom_histogram(binwidth = .1, fill = PAL[1], color = "white") +
  geom_vline(xintercept = median(nd_cat$real), color = PAL[2], linewidth = .9) +
  annotate("text", x = median(nd_cat$real) + .02, y = Inf, vjust = 1.6, hjust = 0,
           label = "typical category: 26%", color = PAL[2], size = 3.4) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "How much of a category's monthly movement is real brand movement",
       y = "Number of categories",
       title = "In most categories, three-quarters of the monthly movement is sampling noise",
       subtitle = "Each bar counts categories. Most sit between 20% and 30% real.")
```

Each bar counts categories, and they pile up between 20% and 30%. In the
typical category about a quarter of the movement is the brand; the rest is the
sample changing.

Two consequences.

- I use **quarterly** figures. Averaging three waves roughly halves the noise.
- There's a hard ceiling on how well *anything* can correlate with monthly
  consideration. The maximum is the square root of the share that's real — a
  quarter real caps you near 0.5, and across these brands it averages 0.48. A
  second identical survey would hit the same ceiling against the first. So a
  mediocre correlation says the ruler is wobbly, not that search is wrong.

The flip side: in small categories search may be the steadier measure. And if
Trends exposed even coarse demographics — Google holds that signal, it's what
powers ad targeting — this comparison would get far more interesting, because
audience cuts are the survey's real advantage.

# 4. Does search rank brands like the funnel does?

For each category I compare brands on average share of search against average
share of consideration and ask whether the orderings agree. Rank agreement
(Spearman's ρ) rather than the shares themselves: the question is whether
search puts brands in the same order, not whether it reproduces the
percentages, which it can't. 1 means identical order, 0 none, negative
reversed.

```{r between, fig.height=6.8}
between <- panel |>
  group_by(category_name, brand_name) |>
  summarise(sos = mean(sos), soc = mean(soc), sop = mean(sop), .groups = "drop")

between_cat <- between |>
  group_by(category_name) |>
  filter(n() >= 4) |>                  # 4+ brands for a rank to mean anything
  summarise(rho = cor(sos, soc, method = "spearman"), .groups = "drop")

ggplot(between_cat, aes(rho, reorder(category_name, rho))) +
  geom_vline(xintercept = 0, color = "grey70") +
  geom_point(color = PAL[1], size = 2.2) +
  scale_x_continuous(limits = c(-1, 1)) +
  labs(x = "Rank agreement between share of search and share of consideration",
       y = NULL,
       title = "Search ranks brands well in some categories and not at all in others",
       subtitle = paste0("Median ", round(median(between_cat$rho), 2),
                         " across ", nrow(between_cat), " categories"))
```

The spread is the finding, so it's worth being precise about both ends.

**At the top**, search reproduces the funnel's order almost exactly: department
stores, online marketplaces, chocolate, streetwear. Categories where the brand
name *is* the search — you type "Kmart" because you're going to Kmart.

**At the bottom**, my first explanation was that nobody searches these
categories. That's wrong, and the real reasons are more useful:

```{r bottom-diagnosis}
# For each weak category, the brand taking the most search, against its
# consideration. A brand pulling far more search than consideration is usually
# the culprit.
between_cat |>
  slice_min(rho, n = 4) |>
  left_join(between |> group_by(category_name) |> slice_max(sos, n = 1),
            by = "category_name") |>
  transmute(category = category_name, rho = round(rho, 2),
            `top search brand` = brand_name,
            `its search` = scales::percent(sos, 1),
            `its consideration` = scales::percent(soc, 1))
```

- **Haircare** isn't unsearched. A brand called *Fruits* takes 65% of category
  search on 3% of consideration — "Fruits" is a word, and my automated matching
  missed it. Same in **Healthy Beverages**: *Shine* and *Remedy* hold 92%
  between them.
- **Men's Online Health** includes a survey option called "Doctors (e.g. GP)"
  with 37% of consideration. It can never have a search share. Survey design,
  not a finding about search.
- **Cafes** fails behaviourally: people find cafes through Maps and "near me",
  not brand names. Krispy Kreme tops search here; Starbucks sits near the
  bottom.

Three of the four are fixable with the tooling that collected the data —
extending my hand-checked entity resolution to the other 42 categories, and
excluding generic survey options from the share. Only Cafes is a real limit.

Skincare settles the "nobody searches this" story: **Skincare 0.77, Body Care
0.75**, but **Female Skincare 0.05** — dragged down by brands called *Software*
and *Go-To*. Same shoppers, opposite scores. Measurement, not consumers.

So the chart mixes three things: categories where search tracks the funnel,
categories where the purchase isn't a branded search, and categories where my
matching failed. The third has an obvious fingerprint, so it's screenable:

```{r contamination-screen}
# A brand holding a large share of search but almost no consideration is
# usually a name collision, not a hidden giant.
flagged <- between |> filter(sos > 0.2, sos > 4 * soc)
flagged |>
  arrange(desc(sos)) |>
  transmute(category = category_name, brand = brand_name,
            search = scales::percent(sos, 1),
            consideration = scales::percent(soc, 1))
```

This needs a human at the end of it. *Adore Beauty* is flagged and isn't an
error — it's an online retailer, so searching the name **is** the shopping
trip. The screen should raise a query, not delete a row.

**How much rides on seven brands?**

```{r contamination-sensitivity}
rank_agreement <- function(d) {
  d |> group_by(category_name) |> filter(n() >= 4) |>
    summarise(rho = cor(sos, soc, method = "spearman"), .groups = "drop")
}
kept <- anti_join(between, flagged, by = c("category_name", "brand_name"))

tibble(version = c("As collected", "7 flagged brands removed"),
       `median rho` = c(round(median(rank_agreement(between)$rho), 2),
                        round(median(rank_agreement(kept)$rho), 2)),
       `categories 0.7+` = c(sum(rank_agreement(between)$rho >= .7),
                             sum(rank_agreement(kept)$rho >= .7)))
```

Seven brands out of four hundred move the median from 0.44 to 0.59. Healthy
Beverages goes −0.30 to 0.40, Garden Care 0.31 to 0.96, Female Skincare 0.05 to
0.57.

The truth sits between the rows: 0.44 understates search because my matching
failed, 0.59 overstates it because dropping a brand also removes a real
competitor from the denominator. Either way the lesson holds — **the quality of
a share-of-search metric is set by the name matching, not the statistics.**

Outliers in *good* categories break the pattern for nameable reasons. Suncorp
and RACQ pull more search than their insurance consideration warrants because
people search them for banking and roadside assistance. The Bottle-O
over-indexes because "bottle-o" is what Australians call any liquor store —
not noise, but a brand that became the generic term, like googling or ubering.
A brand well above the line is worth a conversation, not a correction.

**A parallel from a category not in this sample.** Pet retail isn't one of the
50 here, but it looks like the categories at the top of this chart: a few
well-known chains, distinctive names, a purchase people research online.
Working in that category I saw a retailer whose share of search was falling
while its tracked brand metrics dipped only slightly and its sales stayed flat.
Search moved first, the survey barely registered it, and sales hadn't responded
yet. Nothing was wrong with the survey — it was faithfully recording what
people still said about a brand they were quietly searching for less. That's
the exact gap a search-based early warning is for.



# 5. Does search move with a brand over time?

Ranking brands against each other is static. The harder question is whether one
brand's search and consideration rise and fall together.

```{r within}
# Collapse to quarters, then correlate each brand's own search share against
# its own consideration share over time.
q <- panel |>
  mutate(qtr = as.Date(cut(month, "quarter"))) |>
  group_by(category_name, brand_name, qtr) |>
  summarise(sos = mean(sos), soc = mean(soc), sop = mean(sop), .groups = "drop")

within_brand <- q |>
  group_by(category_name, brand_name) |>
  filter(n() >= 6, sd(sos) > 0, sd(soc) > 0) |>
  summarise(r = cor(sos, soc), .groups = "drop")

ggplot(within_brand, aes(r)) +
  geom_histogram(binwidth = .1, fill = PAL[1], color = "white") +
  geom_vline(xintercept = 0, color = "grey40") +
  labs(x = "How closely a brand's search share and consideration share move together",
       y = "Number of brands",
       title = "Brand by brand, the two barely move together",
       subtitle = paste0("Median ", round(median(within_brand$r), 2),
                         " across ", nrow(within_brand), " brands"))
```

A weak result, and it should be read as one. Some brands track well, more sit
near zero, plenty are negative.

But a correlation this size isn't a conclusion either way. Quarterly survey
figures still carry real noise (§3), and two noisy series can't correlate
strongly however well connected the underlying things are. This rules out
search as a tight substitute. It doesn't say search is uninformative.

Better to stop asking whether the lines move together, and ask what the
distance between them means.

# 6. The Gap/Opportunity Excess Share of Search {#esos}

Marketers already trust a version of this. Binet and Field's excess share of
voice rule: a brand advertising at a bigger share than its market share tends
to grow, one under-investing tends to shrink. The gap predicts direction.
(Summary and sources in [`assumptions.md`](assumptions.md).)

Search gives the same shape of measure without buying media data:

> **Excess Share of Search = share of search − share of consideration.**
> Positive: more people are actively looking for the brand than its
> consideration implies. Negative: the brand is claimed in surveys more than
> it's sought out in real life.

One difference from Binet and Field. Share of voice is an input a brand
controls; share of search is an output of everything it has already done. So
the gap is a symptom, not a lever.

Does today's gap predict where consideration goes over six months?

```{r esos-split}
esos <- q |>
  group_by(category_name, brand_name) |>
  mutate(esos = sos - soc, d_soc = lead(soc, 2) - soc, d_sop = lead(sop, 2) - sop) |>
  ungroup() |>
  filter(!is.na(d_soc))

esos |>
  mutate(group = if_else(esos > 0, "Searched more than considered",
                         "Considered more than searched")) |>
  group_by(group) |>
  summarise(`brand-quarters` = n(),
            `change in consideration share, 6 months later` =
              scales::percent(mean(d_soc), .01), .groups = "drop")
```

```{r esos-chart, fig.height=4.6}
esos |>
  mutate(bin = cut(esos, breaks = quantile(esos, seq(0, 1, .1)),
                   include.lowest = TRUE, labels = FALSE)) |>
  group_by(bin) |>
  summarise(esos = mean(esos), d = mean(d_soc), se = sd(d_soc) / sqrt(n())) |>
  ggplot(aes(esos, d)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_errorbar(aes(ymin = d - 1.96 * se, ymax = d + 1.96 * se), width = 0,
                color = "grey75") +
  geom_point(color = PAL[1], size = 2.8) +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Excess share of search today (brands split into ten equal groups)",
       y = "Change in share of consideration six months later",
       title = "Brands searched for more than they're considered go on to gain consideration",
       subtitle = "Each point is a tenth of the sample; bars show the range the average could sit in")
```

The direction holds across the range: biggest search surplus gains most,
biggest deficit loses most. It isn't a clean straight line, and the two end
groups sit furthest from it — but they follow the same trend, so this isn't one
extreme group dragging the result.

**Does the gap also predict preference?** It should, if search picks up
something real rather than a quirk of the consideration question. It does,
slightly more strongly:

```{r esos-funnel}
tibble(stage = c("Consideration", "Preference"),
       effect = c(coef(lm(d_soc ~ esos + soc, data = esos))[["esos"]],
                  coef(lm(d_sop ~ esos + sop, data = esos))[["esos"]]),
       `std errors from 0` = c(
         summary(lm(d_soc ~ esos + soc, data = esos))$coefficients["esos", 3],
         summary(lm(d_sop ~ esos + sop, data = esos))$coefficients["esos", 3])) |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

A search surplus today shows up at both stages six months later. The funnel
behaving as you'd hope: people search, then consider, then prefer.

**The obvious objection.** Share of consideration mean-reverts — a brand
sitting high has more room to fall — and the gap is *mechanically* negative
when consideration is high, so some of this could be regression to the mean in
a costume. A second version: shares are zero-sum within a category, so one
brand's gain is a rival's loss — competitive by design, but it means a brand
can look like it's winning because a rival collapsed. Controlling for starting
level and category:

```{r esos-models}
# Plain regressions with controls, not a marketing mix model — that's §7.
models <- list(
  "Gap alone"                   = lm(d_soc ~ esos, data = esos),
  "+ starting level"            = lm(d_soc ~ esos + soc, data = esos),
  "+ starting level + category" = lm(d_soc ~ esos + soc + factor(category_name),
                                     data = esos))
purrr::imap_dfr(models, ~ tibble(
  model = .y, effect = coef(.x)[["esos"]],
  `std errors from 0` = summary(.x)$coefficients["esos", 3])) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))
```

Controlling for starting level shrinks the effect — so part of it *was* mean
reversion — but it survives comfortably.

**What the numbers mean.** The effect column is share points gained per share
point of gap; the bottom row, 0.019, is the one to trust. Concretely: **a brand
sitting 10 points of search share ahead of its consideration share gains about
1/5 of a point of consideration share over six months** — 20% heads for 20.2%.

Small, and deliberately so. It's a direction indicator, not a forecast: "which
way is this brand pointing", not "what number will we print". The gap has to be
large and persistent to mean much, which is how excess share of voice is used
in media planning.

This lines up with Tracksuit's own [Return on
Awareness](https://www.gotracksuit.com/report/return-on-awareness) work with
Google, which finds share of search running ahead of market share by 6–12
months. Mine is the same story one step earlier: search runs ahead of
consideration, which runs ahead of share.

**Does this work for small brands?** Commercially the key question — small
brands have the thinnest survey samples. My first cut said no: split by
awareness, the gap looked twice as strong for large brands. That didn't survive
checking. The flagged brands from §4 are small on consideration and huge on
search by construction, so they sit in the small-brand half and drag it down.

```{r esos-by-size}
sized <- esos |>
  left_join(panel |> group_by(category_name, brand_name) |>
              summarise(awareness = mean(awareness), .groups = "drop"),
            by = c("category_name", "brand_name"))
cut_at <- median(sized$awareness, na.rm = TRUE)
eff <- function(d) round(coef(lm(d_soc ~ esos + soc, data = d))[["esos"]], 3)
small <- filter(sized, awareness < cut_at); large <- filter(sized, awareness >= cut_at)

tibble(brands = c("Smaller", "Larger"),
       `as collected` = c(eff(small), eff(large)),
       `flagged removed` = c(
         eff(anti_join(small, flagged, by = c("category_name", "brand_name"))),
         eff(anti_join(large, flagged, by = c("category_name", "brand_name")))))
```

With those seven out, small and large brands behave almost identically. The
apparent weakness was my measurement error, not a property of small brands —
better news commercially, and a better fit with Return on Awareness, which
finds the awareness-to-search link strongest among low and mid-awareness
brands.

**The product version.** One flag on the brand dashboard — "search demand
running ahead of consideration" — plus a quadrant chart placing every brand in
the category on search surplus against consideration. The step after that is
translating the gap into dolalr terms via a marketing mix model, which is where a CMO can defend or ask for the right investment levels inboard meetings land. which is the ultimate aim for Tracksuit 


# 7. How long does search take to show up?

If search leads consideration, by how long? Marketing mix models answer this
with **carryover**: activity doesn't act all at once and doesn't stop dead, it
fades. I use the simplest version — carry some share of last quarter's search
into this quarter, try every share, keep whichever fits best.

```{r carryover, fig.height=4.4}
# Standard adstock:
#   effective search this quarter = (1-k) * this quarter + k * last quarter's
# k is the share carried over. The (1-k) weight matters: without it, higher k
# just inflates the numbers and "improves" the fit for arithmetic reasons.
adstock <- function(x, k) {
  out <- numeric(length(x)); out[1] <- x[1]
  for (i in seq_along(x)[-1]) out[i] <- (1 - k) * x[i] + k * out[i - 1]
  out
}

carry <- purrr::map_dfr(seq(0, 0.9, by = 0.1), function(k) {
  d <- q |>
    group_by(category_name, brand_name) |>
    filter(n() >= 6) |>
    mutate(search = as.numeric(scale(adstock(sos, k))),      # standardise within
           consideration = as.numeric(scale(soc))) |>        # brand so all count equally
    ungroup() |>
    filter(is.finite(search), is.finite(consideration))
  fit <- lm(consideration ~ search, data = d)
  tibble(carryover = k, strength = coef(fit)[["search"]], fit = summary(fit)$r.squared)
})
best <- carry |> slice_max(fit, n = 1)

ggplot(carry, aes(carryover, strength)) +
  geom_line(color = PAL[1], linewidth = .9) +
  geom_point(data = best, color = PAL[2], size = 3.6) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "How much of last quarter's search still counts this quarter",
       y = "Strength of the link to consideration",
       title = "Search feeds into consideration over roughly a quarter",
       subtitle = "Strongest when about half of last quarter's search still counts (orange point)")
```

The best fit carries about half of last quarter's search into this one. In
practice: **a surge in search does most of its work on consideration within
about three months and has largely faded by six.** That matches a separate,
cruder check — shifting the search series forward two to three months is what
maximises its agreement with consideration.

Two limits. The variation explained is small, so the curve gives the *timing*
far more confidently than the *size*. And with 14 quarters per brand this is
near the edge of what the data can identify. A full marketing mix model — media
spend, saturation curves, holdout testing — is the right tool for sizing it,
and well beyond this scope.

**Does this contradict Return on Awareness?** No. They find search leading
*market share* by 6–12 months; I find it leading *consideration* by about
three. Consideration is earlier and faster-moving than sales, so a shorter lead
is what you'd expect. Together they describe one chain: search moves first,
consideration follows within a quarter, market share follows over the year.

# 8. Does it hold up?

The obvious way this could be an illusion: if everyone searches ice cream in
summer *and* considers it in summer, every brand looks connected without any
brand-level signal.

So I broke the brand link deliberately. Each brand keeps its own consideration
history but gets a *rival's* search history from the same category — keeping
category seasonality, destroying the brand connection. A hundred runs gives the
range of results the data produces when there's nothing real to find.

```{r placebo, fig.height=4.2}
# Grey band: 95% range across 100 brand-shuffled runs. Blue: the real result at
# each shift of the search series.
plc <- out("leadlag_placebo.csv")
band <- plc |> filter(run != "real") |> group_by(lag_months) |>
  summarise(lo = quantile(mean_r, .025), hi = quantile(mean_r, .975))
ggplot() +
  geom_ribbon(data = band, aes(lag_months, ymin = lo, ymax = hi), fill = "grey85") +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_line(data = filter(plc, run == "real"), aes(lag_months, mean_r),
            color = PAL[1], linewidth = 1) +
  geom_point(data = filter(plc, run == "real"), aes(lag_months, mean_r),
             color = PAL[1], size = 2.2) +
  annotate("text", x = -3, y = max(band$hi), hjust = 0, vjust = -0.5, size = 3.3,
           color = "grey45", label = "range when the brand link is broken") +
  scale_x_continuous(breaks = -3:3) +
  labs(x = "Search series shifted by k months (positive = search comes first)",
       y = "Average agreement with consideration",
       title = "The brand-level link is real; the exact lead isn't pinned down",
       subtitle = "Blue: real result. Grey: 100 runs with each brand given a rival's search history.")
```

Two claims here, with different strength.

**The brand-level signal is real.** The blue line sits above the grey band at
every shift. If this were only category seasonality, swapping in a rival's
search history would cost nothing and the blue line would sit inside the band.
It doesn't. Supported.

**The 2–3 month lead is not proven.** The line does peak at a two-to-three
month lead, but about one in seven shuffles produces a tilt that large. So:
direction supported, length unproven.

One limitation: the shuffle stays inside the category, so if two brands
genuinely move together the placebo inherits some of that. A conservative test
— which is why I'm comfortable with the first claim and cautious on the
second.

**Other checks.** Raising the search-volume bar doesn't change the ranking
result (0.44 to 0.47 as the cut-off rises). The gap result holds within
categories as well as across them, and at both funnel stages. Every brand whose
search identifier came back empty was automatically re-collected under its
plain name, so "no volume" means no demand rather than a lookup failure.

**Next, in priority order:** test the prediction on later waves it hasn't seen;
fix the name matching in the weak categories and re-run §4; check whether the
gap predicts preference *conversion* rather than preference share. If I were
picking these up at Tracksuit, the holdout test comes first — it's the one that
decides whether this ships.

# 9. What I'd recommend

If I join Tracksuit, this is where I'd start.

**Ship search as a directional layer, not a replacement.** The survey says
where a client stands; the gap says which way they're heading between waves.
That's an addition to a tracker, not a cheaper version of one, and it answers a
real frustration — waiting a month for a number that mostly moves on noise.

**Where it fills gaps.** It fills *time* well: continuous, covering the months
between waves and flagging movement before the next wave confirms it. On
*brands* the answer improved once I fixed the contamination — it works about as
well for small brands as large ones, so this isn't a big-brands-only tool.

The real brand-level limit is a reporting threshold, not missing demand. Trends
rounds to 0–100 within each comparison, so a brand competing against a much
larger rival reads as zero even with real, steady search. "Nobody searches this
brand" and "this brand is too small to show up next to Bunnings" are different
claims and public data can't separate them. Two things would: pulling small
brands against each other rather than the category leader, and the granular
volumes a partnership feed provides. So 1 in 5 is an upper bound on genuine
invisibility, not a measurement of it.

**On qualifying categories.** A category qualifies when most of its brands have
measurable search and search reproduces the funnel's ranking. About a third
pass today, and a chunk of the rest fail on name matching rather than consumer
behaviour — so that share rises with better data engineering.

I'd ship this narrow and early rather than wait for every category to work.
A qualification gate is friction, but cheap friction — a query over data
Tracksuit already holds — and it buys the freedom to put the metric in front of
clients now. Shipping is also how the marketing science improves: a client who
sees their gap turn negative, acts, and watches the next two waves is a natural
experiment I can't run in a notebook. Ten of those beat another round of my own
robustness checks.

What needs protecting is the failure mode. Breaking things is fine when being
wrong costs a shrug; it's expensive when a client opens the dashboard and sees
a brand called *Fruits* leading haircare, because they then distrust every
number on the page. So: ship fast where it passes, label it directional, keep
the contamination screen running.

**What search can't do.** No funnel stages. No absolute levels — a relative
index never says "23% of the category considers you". No audience cuts, at
least from public data; Google holds demographic and location signal
internally, so a partnership feed may differ. Those are the survey's
advantages.

**Where this fits.** Tracksuit already has a Google partnership behind Return
on Awareness. Most limits here — relative indices, sampling wobble,
unidentifiable brands — come from using free public data, not search itself. So
the useful part isn't my numbers. It's the method: how to decide where a search
metric can be trusted, how to check it isn't measuring the wrong thing, and how
to test whether the effect survives the obvious objections.

**Packaging.** The directional flag belongs in the base tracker: it makes the
product stickier between waves and costs almost nothing to produce. A higher
tier is justified by the engineering behind a trustworthy version — curated
term matching for a client's own competitive set, category qualification,
alerting. Charging for the flag means charging for something nearly free;
charging for the work that makes it reliable is defensible. That work is also
getting cheaper: most of the matching here was AI-assisted and took an
afternoon, so I'd measure what a curated competitive set costs to produce and
maintain *now* before pricing it.

# Appendix A: getting the data {-}

Google Trends compares at most five terms per request and scales the index
within that request only. To put every brand in a category on one scale, the
best-known brand rides along in every request as an anchor and each batch is
rescaled to it.

The harder problem is knowing what to search for. "Koala" is a mattress company
and an animal; "Mecca" is a beauty retailer and a city. Where Google's
knowledge graph has an entry for the brand I use that instead of the words,
with hand-written rules for the rest and every decision logged.

I hand-checked terms for eight categories and let automation handle the rest —
exactly where §4's weakest categories came from, and the first thing I'd fix.

Re-running everything on raw brand names with no cleaning barely changes the
category-level answers: median rank agreement 0.55 uncurated versus 0.53
curated, across the 44 categories where both have enough brands. Rank
comparisons forgive a few misplaced brands. At brand level the errors are wild
in both directions — uncurated, "Ultimate" yoghurt looks thousands of times
bigger than it is, while Kärcher and Roll'd vanish over an umlaut and an
apostrophe.

So cleaning barely matters for category research, and is essential for anything
a client sees with their own brand's name on it.

# Appendix B: method notes {-}

- **Share of search** = a brand's index ÷ the category total, monthly, among
  brands clearing the volume bar.
- **Margin of error and real movement.** Expected sampling variance per brand
  is `mean(p(1-p)/n)` across waves, where `n` is the weight of respondents who
  qualified for the category. Real movement is what's left of observed variance
  after subtracting it. Maximum correlation against a perfect predictor is the
  square root of that share. Weights are treated as respondent counts, which
  flatters the survey.
- **Carryover** uses the normalised form `(1-k)·search + k·previous`. The
  unnormalised version in most MMM material inflates the series as `k` rises,
  so a search over `k` picks whichever value manufactures the strongest trend.
  Worth knowing before trusting an off-the-shelf implementation.
- **Assumptions and theory:** [`assumptions.md`](assumptions.md).
- **Reproducing:** `Rscript run_all.R` rebuilds every number and this document
  from the committed raw data. Versions in `session-info.txt`.
