---
title: "Share of Search as a complement to brand tracking"
subtitle: "A pilot study on 50 Australian categories — Tracksuit DS take-home"
author: "Zen"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 5)
source(file.path("R", "functions.R"))

# dataviz palette (fixed categorical order; sequential = single blue hue)
PAL <- c("#2a78d6", "#eb6834", "#1baf7a", "#eda100", "#e87ba4",
         "#008300", "#4a3aa7", "#e34948")
theme_set(theme_minimal(base_size = 12) +
          theme(panel.grid.minor = element_blank(),
                panel.grid.major = element_line(color = "grey92"),
                plot.title.position = "plot"))

survey <- load_survey() |> add_survey_shares()
trends <- load_trends("trends")
trends_naive <- load_trends("trends.naive")
has_search <- !is.null(trends)
has_naive <- !is.null(trends_naive)
PILOTS <- c("Fast Food", "Car Insurance", "Health & Beauty Retailers",
            "Furniture & Homeware", "Mattresses, beds, and pillows",
            "Chocolate", "Liquor Retailer", "Yoghurt")
VOL_MIN <- 0.5   # mean stitched index below this = no measurable search volume

# label helper (ggrepel unavailable via apt; plain non-overlapping text)
ggrepel_or_text <- function(lab_df) {
  geom_text(data = lab_df, aes(label = brand_name), size = 2.8,
            vjust = -0.6, color = "grey30", check_overlap = TRUE)
}
```

# Summary for non-technical readers {-}

**What we did.** We collected 3.5 years of Google search interest for every
brand Tracksuit tracks in 50 Australian categories and compared each brand's
*Share of Search* (its slice of all searches for brands in its category)
against the survey funnel metrics — awareness, consideration, preference.

**What we found.**

Search ranks brands the way the survey does — but only in some categories.
In about a third of the categories we tested, Share of Search reproduces the
survey's competitive ranking closely; in another third it is essentially
uninformative. Which is which is predictable in advance, and that
predictability is what makes this usable: we set out measurable criteria for
whether a category qualifies. About one brand in five has no measurable
search demand at all.

Search cannot yet replace the survey for tracking *movement*. Month-to-month
changes in consideration are mostly survey sampling noise at n≈200 — and we
show that this, not search quality, is what caps any correlation. Search may
in fact be the *steadier* signal in low-sample categories.

Search shows a hint of *leading* consideration by two to three months
(suggestive under a placebo test, not yet established) — consistent with
published evidence that Share of Search leads market share by 6–12 months.
If it holds up under further validation, this is the most commercially
interesting property: an early-warning indicator between survey waves.

**What we recommend.** Use search as a *complement*, not a substitute:
launch a Share-of-Search module in categories that pass a measurable
eligibility test, position it for competitive benchmarking and early trend
detection, and do not position it as a replacement for survey-based funnel
tracking — search carries no demographics, no funnel stages, and no absolute
levels.

*(Caveats: Google Trends is a relative, sampled index; brand-to-search-term
mapping requires curation — we quantify how much it matters; Australia only;
one search source.)*

# 1. The question

Share of Search (SoS) is increasingly cited as a cheap proxy for brand
strength. But "does search correlate with our metrics?" is really two
different product questions:

1. **Between brands:** does SoS rank and size brands within a category the
   way our survey does? If yes, search could *substitute* for parts of
   competitive benchmarking.
2. **Within a brand over time:** do movements in SoS track (or lead)
   movements in that brand's consideration? If yes, search could *complement*
   the survey as a higher-frequency trend signal.

These have different bars to clear and different failure modes, so we test
them separately, on the same data.

Search is treated throughout as a proxy for **consideration** (both imply
intent — one behavioral, one claimed), but we also test which funnel metric
search actually resembles most (§6).

# 2. Data

## 2.1 Survey data

```{r survey-desc}
survey |>
  summarise(categories = n_distinct(category_name),
            `brand-category pairs` = n_distinct(paste(category_name, brand_name)),
            waves = n_distinct(wave_date),
            first_wave = min(wave_date), last_wave = max(wave_date))
```

Monthly syndicated tracking, ~200 qualified respondents per category-wave.
Funnel metrics: prompted awareness, consideration, preference.

## 2.2 Search data

Google Trends, Australia, monthly, September 2021 – March 2025, collected by
`pull/pull_trends.py` (see `pull/README.md`):

- Brands are resolved to Google **knowledge-graph topic entities** where a
  credible one exists (the main defense against homonyms — Koala the mattress
  vs the animal, Mecca the retailer vs the city), with hand-curated terms for
  the pilot categories and every resolution logged for review.
- Trends only compares 5 terms at a time on a relative 0–100 scale, so each
  category's highest-awareness brand rides along in every request as an
  **anchor**, and batches are rescaled to the anchor's level — putting all
  brands in a category on one common scale.
- **Share of Search** = a brand's index / sum of all its category's brand
  indices, per month.

```{r search-desc, eval=has_search}
trends |>
  summarise(categories = n_distinct(category_name),
            `brand-category pairs` = n_distinct(paste(category_name, brand_name)),
            months = n_distinct(month))
```

```{r searchable, eval=has_search}
vol <- trends |>
  group_by(category_name, brand_name) |>
  summarise(mean_idx = mean(trends_index, na.rm = TRUE), .groups = "drop")
searchable_share <- mean(vol$mean_idx >= VOL_MIN)
```

**A finding before any correlation:** only `r if (has_search) scales::percent(searchable_share, 1)`
of tracked brands register measurable search volume at all. Two failure modes
are deliberately separated here. *Resolution failure* — a brand whose search
entity can't be identified — is a fixable data problem, and the pipeline
handles it automatically (a topic entity that returns an empty series is
re-pulled as the raw brand name before we conclude anything). What remains is
*genuine absence of search demand*: the long tail of survey-tracked brands
that Australians simply do not search for at measurable volume. That is a
property of the market, not the pipeline — it alone rules out search as a
full survey substitute, and it becomes the first criterion of the
eligibility framework (§8). Brands below the volume floor are excluded from
SoS computations below.

# 3. How much signal is even in the survey? {#noise}

Before correlating anything we ask what a correlation against monthly survey
metrics *could possibly* achieve. A monthly consideration reading rests on
~200 qualified respondents, so it carries sampling error of roughly ±2–3.5pp.
Decomposing each brand's observed monthly variance into sampling noise vs
plausible true signal:

```{r noise}
nd <- noise_decomposition(survey, "consideration")
nd_cat <- nd |> filter(waves >= 12) |>
  group_by(category_name) |>
  summarise(brands = n(), mean_reliability = mean(reliability, na.rm = TRUE),
            mean_max_r = mean(max_attainable_r, na.rm = TRUE), .groups = "drop")

median_se <- survey |> mutate(se = se_prop(consideration, base_n)) |>
  summarise(m = median(se, na.rm = TRUE)) |> pull(m)

ggplot(nd_cat, aes(mean_reliability, reorder(category_name, mean_reliability))) +
  geom_col(fill = PAL[1], width = .75) +
  scale_x_continuous(labels = scales::percent, expand = expansion(mult = c(0, .05))) +
  labs(x = "Share of monthly variance in consideration that is true signal",
       y = NULL,
       title = "Most monthly movement in consideration is sampling noise",
       subtitle = paste0("Reliability = 1 − sampling var / observed var, averaged over brands. ",
                         "Median monthly SE ±", round(median_se * 100, 1), "pp."))
```

For `r scales::percent(mean(nd$reliability < .5, na.rm = TRUE), 1)` of brands,
**less than half** of monthly variance is signal. The observable correlation
against even a *perfect* covariate is capped at √reliability — e.g. ~0.3 for
Car Insurance at monthly grain. Two design consequences:

- All within-brand analysis uses **quarterly (3-month rolling) smoothing**.
- Weak monthly correlations indict the survey's monthly grain, not search.
  This cuts both ways: in low-sample categories, search may be the *less*
  noisy of the two signals.

# 4. Between brands: does SoS rank brands like the survey does? {#between}

For each category we compare brands' average SoS against their average **share
of consideration** (consideration renormalized across the same searchable
brand set, so both shares live on one denominator).

```{r between-run, eval=has_search}
sos <- trends |>
  semi_join(vol |> filter(mean_idx >= VOL_MIN),
            by = c("category_name", "brand_name")) |>
  group_by(category_name, month) |>
  mutate(sos = trends_index / sum(trends_index, na.rm = TRUE)) |>
  ungroup()

joined <- survey |>
  inner_join(sos |> select(category_name, brand_name, month, trends_index, sos),
             by = c("category_name", "brand_name", "wave_date" = "month")) |>
  group_by(category_name, wave_date) |>
  mutate(soc = consideration / sum(consideration, na.rm = TRUE),
         soa = awareness / sum(awareness, na.rm = TRUE)) |>
  ungroup() |>
  mutate(conversion = if_else(awareness > 0, consideration / awareness, NA_real_))

between_brand <- joined |>
  group_by(category_name, brand_name) |>
  summarise(across(c(sos, soc, soa, consideration, awareness, conversion),
                   ~ mean(.x, na.rm = TRUE)), .groups = "drop")

between_cat <- between_brand |>
  group_by(category_name) |>
  filter(n() >= 4) |>
  summarise(n_brands = n(),
            rho_soc = cor(sos, soc, method = "spearman"),
            rho_aware = cor(sos, soa, method = "spearman"),
            rho_conv = cor(sos, conversion, method = "spearman", use = "complete.obs"),
            .groups = "drop")
```

```{r between-chart, eval=has_search, fig.height=7.5}
ggplot(between_cat,
       aes(rho_soc, reorder(category_name, rho_soc),
           color = category_name %in% PILOTS)) +
  geom_vline(xintercept = 0, color = "grey70") +
  geom_point(size = 2.6) +
  scale_color_manual(values = c(`FALSE` = "grey55", `TRUE` = PAL[2]),
                     labels = c(`FALSE` = "Other", `TRUE` = "Pilot (curated terms)"),
                     name = NULL) +
  scale_x_continuous(limits = c(-1, 1)) +
  labs(x = "Spearman rank correlation: mean SoS vs mean share of consideration",
       y = NULL,
       title = "Share of Search reproduces the survey's brand ranking in most categories",
       subtitle = paste0("Median ρ = ",
                         round(median(between_cat$rho_soc, na.rm = TRUE), 2),
                         "; categories with ≥4 searchable brands"))
```

```{r between-scatter, eval=has_search, fig.height=6}
pb <- between_brand |> filter(category_name %in% PILOTS)
lab <- pb |> group_by(category_name) |>
  mutate(resid = abs(rank(sos) - rank(soc))) |>
  slice_max(resid, n = 2)
ggplot(pb, aes(soc, sos)) +
  geom_abline(color = "grey70", linetype = 2) +
  geom_point(color = PAL[1], size = 2) +
  ggrepel_or_text(lab) +
  facet_wrap(~category_name, scales = "free", ncol = 4) +
  scale_x_continuous(labels = scales::percent_format(1)) +
  scale_y_continuous(labels = scales::percent_format(1)) +
  labs(x = "Share of consideration (survey)", y = "Share of Search",
       title = "Pilot categories: brand-level agreement and interpretable outliers",
       subtitle = "Dashed line = perfect share agreement. Labeled points are the largest rank disagreements.")
```

Interpretation:

- **Agreement is category-specific, not general.** The median category sits
  near ρ = 0.4, but the distribution is wide: roughly a third of categories
  clear ρ ≥ 0.7 (Department Stores, Fast Food, Car Insurance, Online
  Marketplaces — large brands, distinct names, real search demand), while a
  tail sits at or below zero (Haircare, Men's Online Health Providers,
  Healthy Beverages — thin, niche, own-brand-heavy fields where most brands
  barely register). This spread *is* the finding: it is what makes an
  eligibility framework (§8) necessary rather than a blanket rollout.
- **The disagreements are diagnosable, not random.** Suncorp and RACQ
  over-index in search because people search them for banking and roadside
  assistance, not car insurance (intent mix); The Bottle-O over-indexes
  because "bottle-o" is Australian slang for any liquor store. A brand's
  residual from this line is itself interpretable, which is a feature: it
  flags where search intent and brand intent diverge.
- A methodological note worth stating plainly. An earlier iteration of this
  analysis reported a *higher* median ρ (0.62) — because a data-quality bug
  had silently zeroed out ~90 brands, which the volume floor then excluded.
  Fixing the bug added those brands back, and they are precisely the small,
  hard-to-rank ones. **Cleaner data produced a weaker headline number.**
  That is the correct direction of travel: the earlier figure was a
  selection artifact, and anyone benchmarking SoS should ask what was
  dropped before believing a high correlation.

# 5. Within brands: does SoS track movement over time? {#within}

Quarterly-smoothed SoS vs share of consideration, within each brand.

```{r within-run, eval=has_search}
q <- joined |>
  smooth_quarterly(c("sos", "soc")) |>
  filter(!is.na(sos_q), !is.na(soc_q))

within_brand <- q |>
  group_by(category_name, brand_name) |>
  filter(n() >= 10, sd(sos_q) > 0, sd(soc_q) > 0) |>
  summarise(r_within = cor(sos_q, soc_q), .groups = "drop") |>
  left_join(nd |> select(category_name, brand_name, max_attainable_r),
            by = c("category_name", "brand_name"))
```

```{r within-chart, eval=has_search}
ggplot(within_brand, aes(r_within)) +
  geom_histogram(binwidth = .1, fill = PAL[1], color = "white") +
  geom_vline(xintercept = 0, color = "grey40") +
  geom_vline(aes(xintercept = median(r_within)), color = PAL[2], linewidth = .8) +
  annotate("text", x = median(within_brand$r_within) + .03, y = Inf,
           label = paste0("median = ", round(median(within_brand$r_within), 2)),
           hjust = 0, vjust = 1.5, color = PAL[2]) +
  labs(x = "Within-brand correlation (quarterly SoS vs share of consideration)",
       y = "Brands",
       title = "Within-brand tracking is weak on average — but the ceiling is low too",
       subtitle = paste0("Mean attainable ceiling from survey noise ≈ ",
                         round(mean(within_brand$max_attainable_r, na.rm = TRUE), 2),
                         " even for a perfect covariate"))
```

The distribution centers near zero with wide spread. Judged naively this
kills the trend-tracking use case; judged against §3's noise ceiling it is
*inconclusive at this sample size* — the survey's monthly grain cannot
falsify or confirm a modest true relationship over 3.5 years. What the data
*can* say:

- A minority of brands (large, dynamic ones — e.g. Youi) show strong
  agreement; these are exactly the brands with the most true signal in both
  series.
- Regressions on within-brand standardized series (each brand's SoS and
  share-of-consideration centered and scaled before pooling — the demeaning
  step of a brand fixed-effects estimator) give small positive coefficients
  in most categories — directionally right, small, and attenuated by noise
  in both variables.

## 5.1 Does search *lead* consideration?

```{r leadlag, eval=has_search}
leadlag <- purrr::map_dfr(-3:3, function(k) {
  q |>
    group_by(category_name, brand_name) |>
    arrange(wave_date, .by_group = TRUE) |>
    mutate(sos_l = if (k >= 0) lag(sos_q, k) else lead(sos_q, -k)) |>
    filter(!is.na(sos_l), !is.na(soc_q)) |>
    group_by(category_name, brand_name) |>
    filter(n() >= 10, sd(sos_l) > 0, sd(soc_q) > 0) |>
    summarise(r = cor(sos_l, soc_q), .groups = "drop") |>
    summarise(lag_months = k, mean_r = mean(r), n_brands = n())
})

ggplot(leadlag, aes(lag_months, mean_r)) +
  geom_col(fill = ifelse(leadlag$lag_months > 0, PAL[1], "grey70"), width = .7) +
  scale_x_continuous(breaks = -3:3,
                     labels = c("-3", "-2", "-1", "0", "+1", "+2", "+3")) +
  labs(x = "SoS shifted by k months (positive = search earlier than survey)",
       y = "Mean within-brand correlation",
       title = "Search interest weakly leads surveyed consideration",
       subtitle = "Correlation is highest when SoS is shifted 2–3 months ahead of the survey")
```

The asymmetry is small but consistent: SoS from 2–3 months ago correlates
better with today's consideration than contemporaneous SoS, and better than
the reverse direction.

**Placebo check.** To test whether any of this is brand-level signal rather
than category seasonality, we reassign each SoS series to a random *other*
brand in the same category (100 permutations) and recompute the whole curve
(`analysis/03_leadlag_placebo.R`):

```{r placebo, eval=file.exists("out/leadlag_placebo.csv")}
plc <- readr::read_csv("out/leadlag_placebo.csv", show_col_types = FALSE)
band <- plc |> filter(run != "real") |> group_by(lag_months) |>
  summarise(lo = quantile(mean_r, .025), hi = quantile(mean_r, .975))
real_c <- plc |> filter(run == "real")
ggplot() +
  geom_ribbon(data = band, aes(lag_months, ymin = lo, ymax = hi),
              fill = "grey85") +
  annotate("text", x = -3, y = max(band$hi), hjust = 0, vjust = -0.4,
           label = "placebo range", color = "grey45", size = 3.2) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_line(data = real_c, aes(lag_months, mean_r), color = PAL[1], linewidth = 1) +
  geom_point(data = real_c, aes(lag_months, mean_r), color = PAL[1], size = 2.4) +
  scale_x_continuous(breaks = -3:3) +
  labs(x = "SoS shifted by k months (positive = search earlier than survey)",
       y = "Mean within-brand correlation",
       title = "The brand-level link is real; the lead asymmetry is suggestive",
       subtitle = "Grey band = 95% range of 100 within-category brand-shuffled placebos")
```

Two conclusions, at different confidence levels. The **brand-level
relationship is real**: the true curve sits above the placebo band across
lags, so SoS carries brand-specific information about consideration beyond
category-wide seasonality. The **lead asymmetry is suggestive but not
established**: the real +2-vs−2 asymmetry (+0.033) is exceeded by 14 of 100
placebo permutations (p ≈ 0.14). We report it as the pattern you'd expect if
search precedes claimed consideration — and note it is directionally
consistent with Les Binet's and Tracksuit's own published finding that Share
of Search leads *market share* by 6–12 months (see §6). Validating this
properly (more waves, holdout prediction) is the single highest-value next
step, because a confirmed lead turns SoS into an early-warning metric the
survey cannot provide at monthly cost. One caveat applies to any version of
this analysis, theirs and ours: media campaigns move awareness, search, and
consideration together on different response lags, so lead-lag alone cannot
establish causation.

# 6. Which funnel metric does search actually reflect? {#funnel}

```{r funnel, eval=has_search}
funnel_cmp <- between_cat |>
  select(category_name, Consideration = rho_soc, Awareness = rho_aware,
         `Consideration/Awareness` = rho_conv) |>
  pivot_longer(-category_name, names_to = "metric", values_to = "rho")

ggplot(funnel_cmp, aes(rho, metric, color = metric)) +
  geom_vline(xintercept = 0, color = "grey70") +
  geom_jitter(height = .12, alpha = .5, size = 1.8) +
  stat_summary(fun = median, geom = "point", shape = 18, size = 5) +
  scale_color_manual(values = PAL[c(1, 2, 3)], guide = "none") +
  labs(x = "Between-brand Spearman ρ vs SoS (one point per category)",
       y = NULL,
       title = "Search resembles awareness as much as consideration",
       subtitle = "Large diamonds = median across categories")
```

Between brands, SoS correlates with awareness about as strongly as with
consideration — funnel shares are heavily collinear across brands (big brands
are big everywhere). The sharper test is *conversion* (consideration ÷
awareness): if SoS only reflected salience, it would not track how well
awareness converts to intent. The conversion correlation is positive but
weaker — consistent with search sitting *between* salience and intent.

This is not an inconvenient result — it independently replicates Tracksuit's
own published position. The Tracksuit × Google [*Return on Awareness*
study](https://www.gotracksuit.com/report/return-on-awareness) (31 brands, 15
AU/NZ categories, 2024–25) links **awareness** to Share of Search (~+5pts SoS
per +5pts awareness around the 30%-awareness mark, strongest for low/mid
awareness brands) and positions SoS as a leading indicator of market share.
Our 50-category, 3.5-year study on public data arrives at the same
qualitative conclusion from the other direction: SoS behaves as a general
**brand-demand** signal spanning awareness and consideration, not a
one-to-one consideration proxy. Practical consequence: present SoS within
the awareness → SoS → sales diagnostic stack that study proposes, rather
than as a substitute for any single funnel metric. (Their finding of
diminishing returns at high awareness also cautions against the linear-share
lens for mature brands; a logit-scale sensitivity left for follow-up.)

# 7. Validation {#validation}

## 7.1 Measurement-error robustness: curated vs uncurated terms

```{r curation, eval=has_naive}
naive_vol <- trends_naive |>
  group_by(category_name, brand_name) |>
  summarise(mean_idx = mean(trends_index, na.rm = TRUE), .groups = "drop")
sos_n <- trends_naive |>
  semi_join(naive_vol |> filter(mean_idx >= VOL_MIN),
            by = c("category_name", "brand_name")) |>
  group_by(category_name, month) |>
  mutate(sos = trends_index / sum(trends_index, na.rm = TRUE)) |>
  ungroup()
joined_n <- survey |>
  inner_join(sos_n |> select(category_name, brand_name, month, sos),
             by = c("category_name", "brand_name", "wave_date" = "month")) |>
  group_by(category_name, wave_date) |>
  mutate(soc = consideration / sum(consideration, na.rm = TRUE)) |>
  ungroup()
between_n <- joined_n |>
  group_by(category_name, brand_name) |>
  summarise(sos = mean(sos), soc = mean(soc), .groups = "drop") |>
  group_by(category_name) |> filter(n() >= 4) |>
  summarise(rho_naive = cor(sos, soc, method = "spearman"), .groups = "drop")

cmp <- between_cat |> select(category_name, rho_curated = rho_soc) |>
  inner_join(between_n, by = "category_name")

ggplot(cmp, aes(rho_naive, rho_curated)) +
  geom_abline(color = "grey70", linetype = 2) +
  geom_point(color = PAL[1], size = 2.4) +
  coord_equal(xlim = c(-1, 1), ylim = c(-1, 1)) +
  labs(x = "Uncurated (raw brand names)", y = "Curated (entities + disambiguation)",
       title = "Curation does not systematically improve category-level correlation",
       subtitle = "Between-brand ρ per category; scatter around the line is large in both directions")
```

```{r curation-placeholder, eval=!has_naive, echo=FALSE, results='asis'}
cat("> *The uncurated-tier pull has not landed yet; this section will populate",
    "automatically when `data/raw/trends.naive/` exists.*\n")
```

Search-term choice is the largest source of measurement error on the search
side, so we re-pull all 507 brands as raw names with no entity resolution and
no curation (the `--naive` flag) and repeat the analysis. This tests whether
our conclusions are artifacts of our particular term choices.

**The result contradicted our expectation, and that is worth reporting
plainly.** We expected curation to visibly lift category-level correlations.
It does not: median between-brand ρ is 0.53 curated vs 0.55 uncurated, with
curation ahead in 23 of 44 categories and a mean gain of +0.003 — a coin
flip.

Yet at the *brand* level the two tiers disagree violently, in both
directions:

```{r curation-brand, eval=has_naive}
cv <- trends |> group_by(category_name, brand_name) |>
  summarise(curated = mean(trends_index), .groups = "drop")
nv <- trends_naive |> group_by(category_name, brand_name) |>
  summarise(uncurated = mean(trends_index), .groups = "drop")
cmp_b <- inner_join(cv, nv, by = c("category_name", "brand_name")) |>
  mutate(ratio = (uncurated + .01) / (curated + .01))
bind_rows(
  cmp_b |> slice_max(ratio, n = 5) |>
    mutate(failure = "Uncurated inflated (homonym)"),
  cmp_b |> slice_min(ratio, n = 5) |>
    mutate(failure = "Uncurated missed the brand (punctuation/accent)")
) |>
  select(failure, category_name, brand_name, curated, uncurated) |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

Two distinct failure modes, each invisible in the aggregate. Uncurated terms
*inflate* brands whose names are common words — "Ultimate" yoghurt picks up
3,000× its real volume, "Software" skincare 1,900×, "Eva" mattresses, "Grill'd"
fast food. They also *erase* brands whose literal survey spelling is not how
people type: "Roll’d" (curly apostrophe), "Kärcher" (umlaut), "Bed Bath N’
Table", "Temple & Webster" — all read as zero uncurated, all substantial
brands in reality.

Why doesn't this move the category correlations? Because Spearman ρ over 8–15
brands is robust to a few misplaced ones, and the two error types partly
offset. The honest conclusion is therefore split by use case:

- **For research conclusions at category level** (this study's claims), the
  findings are robust to term curation. We would have reached the same
  verdict with raw terms.
- **For any brand-level product** — a client dashboard reading "your Share of
  Search is 12%" — curation is non-negotiable. An uncurated pipeline would
  have told a yoghurt brand called Ultimate that it dominated its category,
  and told Kärcher it did not exist.

That distinction is itself the deliverable: it tells Tracksuit where the
data-engineering budget has to go (customer-facing brand metrics) and where
it does not (category screening and research).

## 7.2 Homonym placebo

The failure mode curation prevents, on one brand. Koala is an Australian
furniture and mattress company; it is also an animal. The uncurated pull
takes the raw term and measures the animal:

```{r homonym, eval=has_naive, fig.height=4}
hom <- bind_rows(
  trends |> filter(category_name == "Furniture & Homeware", brand_name == "Koala") |>
    mutate(tier = "Curated (topic entity: Koala Sleep Pty Ltd)"),
  trends_naive |> filter(category_name == "Furniture & Homeware", brand_name == "Koala") |>
    mutate(tier = "Uncurated (raw term: \"Koala\")")
)
ggplot(hom, aes(month, raw_index, color = tier)) +
  geom_line(linewidth = .9) +
  scale_color_manual(values = c(PAL[1], PAL[2]), name = NULL) +
  theme(legend.position = "top") +
  labs(x = NULL, y = "Google Trends index (raw, within batch)",
       title = "The same brand name, two entirely different signals",
       subtitle = "Uncurated 'Koala' tracks tourist and school-holiday interest in the animal")
```

The two series are barely related. Only the curated one is a measurement of
the brand, and the difference is not subtle noise — it is a different subject
entirely. Multiply this by the ~100 brands in this sample carrying a common
noun, a place name, or a first name, and it is the single largest threat to
naive Share-of-Search work.

## 7.3 Robustness

- **Volume floor sensitivity:** the between-brand result is stable across
  floors — median ρ = 0.44 / 0.46 / 0.47 / 0.46 at floors of 0.5 / 1 / 2 / 5
  mean index points (410 / 392 / 342 / 267 brands). Excluding more marginal
  brands does not manufacture a stronger relationship.
- **Anchor stitching:** the pull flags any batch where the anchor brand has
  no volume (`anchor-zero`), since the rescaling is then undefined; the
  final dataset carries zero such flags. Stitch scale factors are logged per
  batch for audit.
- **Entity-resolution failure is detected, not assumed away:** any brand
  whose topic entity returns an empty series is automatically re-pulled as a
  raw term (110 brands in the final data), so "no search volume" means no
  demand rather than a bad identifier. This check was added after QA found
  ~90 well-known brands (Myer, Big W, TikTok, Liquorland) reading as zero
  because Google's autocomplete had offered a dud duplicate entity.
- **Re-pull stability:** Trends indices are sampled; repeated pulls of one
  category on different days shift monthly values by a few index points but
  leave SoS ranks essentially unchanged.
- **Not tested here** (out of timebox, would be next): holdout-wave
  prediction, other geographies, non-Google search sources, demographic
  cuts (Trends has none).

# 8. Recommendation {#recommendation}

```{r eligibility, eval=has_search}
eligibility <- between_cat |>
  left_join(vol |> group_by(category_name) |>
              summarise(pct_searchable = mean(mean_idx >= VOL_MIN)),
            by = "category_name") |>
  left_join(nd_cat |> select(category_name, survey_reliability = mean_reliability),
            by = "category_name") |>
  mutate(eligible = rho_soc >= .7 & pct_searchable >= .6) |>
  arrange(desc(rho_soc))
eligibility |>
  select(category_name, n_brands, rho_soc, pct_searchable, eligible) |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

**1. Launch SoS as a complement, in eligible categories.** A category
qualifies on measurable criteria, not intuition: (a) most of its brands
register search volume; (b) SoS reproduces the survey's brand ranking
(ρ ≥ 0.7 in a validation window); (c) brand names are resolvable to clean
search entities. Roughly a third of the 50 sampled categories pass today.
The product surface: competitive benchmarking between waves, and — pending
further validation — a 2–3 month early-warning signal on consideration.

**2. Do not substitute.** Search has no demographics, no funnel stages, no
absolute levels (a relative index cannot say "23% of the category considers
you"), and a third of tracked brands are invisible to it. The survey remains
the source of truth for levels; search adds frequency and cost-efficiency.

**3. Build on the existing Google partnership; treat public Trends as the
prototype tier.** Tracksuit already works with Google's aggregated search
data (*Return on Awareness*). Most of the limitations this study documents —
relative indices, sampling noise, 5-term batching, invisible long-tail
brands, homonym contamination — are artifacts of the *free public* tier, not
of search data itself. The path is therefore: use this study's methodology
(eligibility criteria, noise-aware validation, entity resolution with an
auditable log) as the blueprint, and the partnership data as the production
source. §7.1 sharpens where that engineering effort belongs: entity
resolution is essential for *brand-level* client reporting and largely
irrelevant for *category-level* screening.

**Worked example — a category not in this sample:** pet retail (a category
I've worked in) would likely score well: distinctive brand names (Petbarn,
PETstock), a search-heavy purchase journey, and high category interest —
the framework predicts eligibility before spending a survey wave to verify.

# Appendix: methods detail {-}

- **Stitching:** with anchor $a$ and batch $k$, each series is scaled by
  $\bar{x}_{a,\text{ref}} / \bar{x}_{a,k}$; SoS is then computed within
  category-month over the searchable brand set.
- **Noise decomposition:** for brand $b$, $\hat\sigma^2_{noise} =
  \overline{p_t(1-p_t)/n_t}$; reliability $= 1 - \hat\sigma^2_{noise} /
  \hat\sigma^2_{obs}$; the attainable correlation against a noiseless
  covariate is $\sqrt{\text{reliability}}$. Weight-based $n_t$ understates
  design effects, so ceilings are if anything optimistic.
- **Reproducibility:** `pull/pull_trends.py --all` (and `--all --naive`)
  regenerates `data/raw/`; `rmarkdown::render("analysis/report.Rmd")`
  rebuilds this document. Raw pull responses and the resolution log are
  committed.
- **Limitations:** single search source; relative indices; AU only; SoS
  denominator limited to tracked brands (a competitor outside the tracked
  set is invisible); intent mix within brand terms (Suncorp banking vs
  insurance) is unresolved without query-level data.
