---
title: "Share of Search and the brand funnel"
subtitle: "Can search data complement or substitute survey-based brand tracking? A pilot study across 50 Australian categories"
author: "Zain Raza"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 4.8)
source(file.path("R", "functions.R"))
PAL <- c("#2a78d6", "#eb6834", "#1baf7a", "#eda100")
theme_set(theme_minimal(base_size = 12) +
          theme(panel.grid.minor = element_blank(),
                panel.grid.major = element_line(color = "grey92"),
                plot.title.position = "plot"))
```

```{r loading}
# A brand needs a minimum average Trends index to be usable. Trends reports
# 0-100 within each request, so a brand averaging under half a point can't
# support a share calculation. Section 8 shows a higher bar changes little.
VOL_MIN <- 0.5

survey <- load_survey()            # 50 categories of funnel metrics
trends <- load_trends("trends")    # Google Trends, collected in pull/

vol <- trends |>
  group_by(category_name, brand_name) |>
  summarise(mean_idx = mean(trends_index, na.rm = TRUE), .groups = "drop")
searchable <- filter(vol, mean_idx >= VOL_MIN)

# The analysis table: keep usable brands, turn search volume into share of
# search within each category-month, join the survey, do the same for the
# funnel metrics so everything is a share of its category.
panel <- trends |>
  semi_join(searchable, by = c("category_name", "brand_name")) |>
  group_by(category_name, month) |>
  mutate(sos = trends_index / sum(trends_index, na.rm = TRUE)) |>
  ungroup() |>
  inner_join(survey, by = c("category_name", "brand_name", "month" = "wave_date")) |>
  group_by(category_name, month) |>
  mutate(soc = consideration / sum(consideration, na.rm = TRUE),
         sop = preference / sum(preference, na.rm = TRUE)) |>
  ungroup()

# Only the 100-shuffle placebo is read from disk — it takes two minutes, so it
# runs in analysis/03_leadlag_placebo.R. Everything else computes here.
out <- function(f) readr::read_csv(file.path("out", f), show_col_types = FALSE)
```

# Summary {-}

Tracksuit's survey asks people what they *say*: heard of this brand, would you
consider it, do you prefer it. Search asks nothing. It just records what people
*do*. I wanted to know how much of the saying the doing can replace.

**Three findings.**

**1. Search ranks brands like the survey does in some categories, not in
others.** So the product isn't a metric, it's a qualification check: which
categories can carry a search-based number before it reaches a client.

**2. Search can't replace the survey for month-to-month tracking — because of
the survey's sample size, not search's quality.** At ~200 respondents a month,
consideration carries a margin of error of about ±3 points. In the typical
category only a quarter of the monthly movement is real. Nothing correlates
well with a number that unsteady.

Scale is the real trade here. Surveys do scale, but every extra brand,
category or country costs sample, and precision comes from sample *per cell*.
Scale out and each number gets shakier, not steadier. Search is the opposite:
adding brands and countries is nearly free, and it never runs out of
respondents. What you give up is control. You can't ask it a question, and you
can't cut it by audience.

**3. The useful signal is the gap between the two, not the correlation between
them.** Brands searched for more than their consideration share implies tend to
gain consideration, and preference, over the next six months. Brands searched
for less tend to lose it.

**Recommendation.** Ship search as a directional early-warning layer on top of
the funnel. The survey says where a brand stands. The gap says which way it's
heading between waves. It won't replace the survey — no demographics, no funnel
stages, no absolute levels. And you can't throw it together: seven badly
matched brand names swung my headline result by 15 points. That's a data
engineering job, not a research one.

> The commercial case isn't cheaper tracking. It's faster tracking.

Clients wait a month for a number that mostly moves on noise. A search-gap flag
gives them a direction between waves, for free. I'd put it in front of clients
in the categories where it holds up and learn from what they do with it, rather
than perfecting every category first. A client who acts on a search signal and
sees the next wave confirm it teaches us more than another month of internal
validation.

# 1. The question, in funnel terms

The funnel is awareness → consideration → preference. Each step is a claim a
respondent makes about themselves.

A search is a behaviour, and it sits between the first two steps. You can't
search a brand you've never heard of, so search needs awareness. And typing a
brand name is more effort than recognising a logo, so it's closer to intent. If
that's right, share of search reads the top of the funnel behaviourally — and
should move before the survey, because people act before they get asked.

Two product questions, different bars:

| Question | If yes, search is | What has to be true |
|---|---|---|
| Does search rank brands like the funnel does? | a benchmarking layer | search ranks brands like consideration |
| Does search move *before* the funnel? | an early-warning layer | today's search carries information about tomorrow's consideration |

Theory and assumptions are in [`assumptions.md`](assumptions.md).

# 2. The data

Google Trends for every brand in all 50 Australian categories, September 2021
to March 2025, matched to the survey waves. **Share of search** is a brand's
index divided by the total for its category, which puts it on the same footing
as share of consideration.

```{r data-summary}
tibble(categories = n_distinct(panel$category_name),
       brands = n_distinct(paste(panel$category_name, panel$brand_name)),
       months = n_distinct(panel$month),
       `brands with too little search` = nrow(vol) - nrow(searchable))
```

One number is a finding on its own. Of the 507 brands tracked here, 97 — about
1 in 5 — never register enough search volume to measure. Read that as an upper
bound rather than a count: Trends rounds to a 0–100 index within each
comparison, so a small brand sitting beside a category giant can read as zero
even with real, steady search (§9 returns to this). Either way the direction is
clear enough to settle the substitution question — you can't run a tracker on
brands you can't see.

Collecting this was most of the work and none of the insight. Google compares
five terms per request and scales the index inside each one, so getting 507
brands onto a single scale takes stitched, rate-limited requests plus a lot of
disambiguation — "Koala" is an animal, "Mecca" is a city, "Ultimate" is a
yoghurt. I automated the collection, entity resolution and QA so this could
cover 50 categories in the time available instead of the five I'd have checked
by hand. The full run takes about 90 minutes, so the raw data is committed —
nobody has to re-run it to check my numbers. Mechanics in Appendix A.

# 3. How steady is the survey itself?

Each brand's monthly score comes from the people who qualified for that
category that month — about 200 of them. For a percentage based on 200 people,
the margin of error is roughly ±3 points. A brand moving from 25% to 28% is
well inside what sampling alone produces.

I can measure this per brand: take how much a brand's consideration actually
moved, subtract how much sampling alone would produce given its score and its
sample size, and what's left is real movement. Here's one brand with its
sampling margin drawn around it:

```{r noise-example, fig.height=4}
eg <- survey |>
  filter(category_name == "Fast Food", brand_name == "KFC") |>
  arrange(wave_date) |>
  mutate(se = se_prop(consideration, base_n),
         q = zoo::rollmean(consideration, 3, fill = NA, align = "right"))

ggplot(eg, aes(wave_date, consideration)) +
  geom_ribbon(aes(ymin = consideration - 1.96 * se, ymax = consideration + 1.96 * se),
              fill = "grey86") +
  geom_line(color = "grey55", linewidth = .5) +
  geom_line(aes(y = q), color = PAL[1], linewidth = 1.1) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Consideration",
       title = "One brand's monthly consideration, with its margin of error",
       subtitle = "Grey: where the score could sit on sampling alone. Blue: the three-month average I use instead.")
```

This isn't a criticism of the survey. The blue line is the best available
estimate and the fieldwork is doing its job. It's about resolution. Ask 200
people and you can see a ten-point gap between brands; you can't see a
two-point move in one brand month to month. Treating every wiggle as real is
what produces false alarms. So I compare quarters, not months.

Across all brands, by category:

```{r noise, fig.height=4.4}
nd <- noise_decomposition(survey, "consideration")
nd_cat <- nd |> filter(waves >= 12) |> group_by(category_name) |>
  summarise(real = mean(reliability, na.rm = TRUE), .groups = "drop")

ggplot(nd_cat, aes(real)) +
  geom_histogram(binwidth = .1, fill = PAL[1], color = "white") +
  geom_vline(xintercept = median(nd_cat$real), color = PAL[2], linewidth = .9) +
  annotate("text", x = median(nd_cat$real) + .02, y = Inf, vjust = 1.6, hjust = 0,
           label = "typical category: 26%", color = PAL[2], size = 3.4) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "How much of a category's monthly movement is real brand movement",
       y = "Number of categories",
       title = "In most categories, three-quarters of the monthly movement is sampling noise",
       subtitle = "Each bar counts categories. Most sit between 20% and 30% real.")
```

Each bar counts categories. They pile up between 20% and 30%. In the typical
category, about a quarter of what you see moving is the brand moving. The rest
is the sample changing.

This has two consequences.

- I use **quarterly** figures. Averaging three waves roughly halves the noise.
- There's a hard ceiling on how well *anything* can correlate with monthly
  consideration. If a quarter of the movement is real, a perfect predictor is
  still chasing a mostly-noise target. The maximum correlation is the square
  root of the share that's real — a quarter real caps you near 0.5, and across
  these brands it averages 0.48. A second identical survey would hit the same
  ceiling against the first. So a mediocre correlation tells you the ruler is
  wobbly, not that search is wrong.

The flip side: in small categories, search may be the steadier measure. And if
Trends exposed even coarse demographics — Google clearly holds that signal,
it's what powers ad targeting — this comparison would get far more interesting,
because audience cuts are the survey's real advantage.

# 4. Does search rank brands like the funnel does?

For each category I compare brands on average share of search against average
share of consideration, and ask whether the orderings agree. I use rank
agreement (Spearman's ρ) rather than comparing the shares directly: the
question is whether search puts brands in the same order, not whether it
reproduces the percentages, which it can't.

1 means identical order, 0 means no relationship, negative means reversed.

```{r between, fig.height=6.8}
between <- panel |>
  group_by(category_name, brand_name) |>
  summarise(sos = mean(sos), soc = mean(soc), sop = mean(sop), .groups = "drop")

between_cat <- between |>
  group_by(category_name) |>
  filter(n() >= 4) |>                  # 4+ brands for a rank to mean anything
  summarise(rho = cor(sos, soc, method = "spearman"), .groups = "drop")

ggplot(between_cat, aes(rho, reorder(category_name, rho))) +
  geom_vline(xintercept = 0, color = "grey70") +
  geom_point(color = PAL[1], size = 2.2) +
  scale_x_continuous(limits = c(-1, 1)) +
  labs(x = "Rank agreement between share of search and share of consideration",
       y = NULL,
       title = "Search ranks brands well in some categories and not at all in others",
       subtitle = paste0("Median ", round(median(between_cat$rho), 2),
                         " across ", nrow(between_cat), " categories"))
```

The spread is the finding, so both ends are worth being precise about.

**At the top**, search gets the order almost exactly right: department stores,
online marketplaces, chocolate, streetwear. Categories where the brand name *is*
the search — you type "Kmart" because you're going to Kmart.

**At the bottom**, my first explanation was that nobody searches these
categories. That's wrong, and the real reasons are more useful:

```{r bottom-diagnosis}
# For each weak category, the brand taking the most search, against its
# consideration. A brand pulling far more search than consideration is usually
# the culprit.
between_cat |>
  slice_min(rho, n = 4) |>
  left_join(between |> group_by(category_name) |> slice_max(sos, n = 1),
            by = "category_name") |>
  transmute(category = category_name, rho = round(rho, 2),
            `top search brand` = brand_name,
            `its search` = scales::percent(sos, 1),
            `its consideration` = scales::percent(soc, 1))
```

- **Haircare** isn't unsearched. A brand called *Fruits* takes 65% of the
  category's search on 3% of consideration. "Fruits" is a word, and my
  automated matching didn't catch it. Same in **Healthy Beverages**, where
  *Shine* and *Remedy* hold 92% between them.
- **Men's Online Health** contains a survey option called "Doctors (e.g. GP)"
  with 37% of consideration. It can never have a brand search share. That's
  survey design, not a finding about search.
- **Cafes** fails for a real behavioural reason: people find cafes through Maps
  and "near me", not by typing brand names. Krispy Kreme tops search here — a
  destination people look up — while Starbucks sits near the bottom.

Three of those four are fixable with the same tooling that collected the data.
The name-matching failures need the entity resolution I applied to my eight
hand-checked categories, extended to the other 42. The generic survey option
needs excluding from the share calculation. Only Cafes is a genuine limit.

Skincare settles the "nobody searches this" story: **Skincare scores 0.77 and
Body Care 0.75**, while **Female Skincare scores 0.05** — dragged down by
brands called *Software* and *Go-To*. These are the same shoppers behaving the
same way in both categories, and the scores come out opposite. The brands
dragging Female Skincare down are ordinary words that Google cannot tell apart
from the brand, so the most likely explanation is my measurement rather than
anything about the consumers.

So this chart mixes three things: categories where search genuinely tracks the
funnel, categories where the purchase isn't a branded search, and categories
where my matching failed. The third group has an obvious fingerprint, which
makes it screenable:

```{r contamination-screen}
# A brand holding a large share of search but almost no consideration is
# usually a name collision, not a hidden giant.
flagged <- between |> filter(sos > 0.2, sos > 4 * soc)
flagged |>
  arrange(desc(sos)) |>
  transmute(category = category_name, brand = brand_name,
            search = scales::percent(sos, 1),
            consideration = scales::percent(soc, 1))
```

This needs a human at the end of it. *Adore Beauty* is flagged and isn't an
error — it's an online retailer, so searching the name **is** the shopping
trip. The screen should raise a query, not delete a row.

**How much rides on seven brands?**

```{r contamination-sensitivity}
rank_agreement <- function(d) {
  d |> group_by(category_name) |> filter(n() >= 4) |>
    summarise(rho = cor(sos, soc, method = "spearman"), .groups = "drop")
}
kept <- anti_join(between, flagged, by = c("category_name", "brand_name"))

tibble(version = c("As collected", "7 flagged brands removed"),
       `median rho` = c(round(median(rank_agreement(between)$rho), 2),
                        round(median(rank_agreement(kept)$rho), 2)),
       `categories 0.7+` = c(sum(rank_agreement(between)$rho >= .7),
                             sum(rank_agreement(kept)$rho >= .7)))
```

Seven brands out of four hundred move the median from 0.44 to 0.59. Healthy
Beverages goes from −0.30 to 0.40, Garden Care from 0.31 to 0.96, Female
Skincare from 0.05 to 0.57.

The truth sits between the two rows — 0.44 understates search because my
matching failed, 0.59 overstates it because dropping a brand also removes a
real competitor from the denominator. Either way, the lesson is the same: **the
quality of a share-of-search metric is set by the name matching, not the
statistics.**

The brands that break the pattern in *good* categories break it for nameable
reasons. Suncorp and RACQ pull more search than their insurance consideration
warrants, because people search them for banking and roadside assistance. The
Bottle-O over-indexes because "bottle-o" is what Australians call any liquor
store — which isn't noise, it's a brand that became the generic term, like
googling or ubering. A brand well above the line is worth a conversation, not a
correction.

**A parallel from a category not in this sample.** Pet retail isn't one of the
50 here, but it looks like the categories at the top of this chart: a few
well-known chains, distinctive names, a purchase people research online. I
built a marketing mix model for a pet retail brand that was holding sales flat
through promotional depth while quietly losing market share and share of voice.
The sales line looked fine; the brand indicators underneath it were eroding,
and the promotions masked that in the only number the business was watching.

I used share of voice there as the leading indicator, and that is the same
argument this study makes with search — share of search is the version you can
get for every brand in a category without buying media data. It updates
monthly, costs nothing, and doesn't depend on a competitor's spend being
visible to you. That's the whole appeal: the early-warning logic already works,
this is a cheaper way to observe it.

# 5. Does search move with a brand over time?

Ranking brands against each other is static. The harder question is whether one
brand's search and consideration rise and fall together.

```{r within}
# Collapse to quarters, then correlate each brand's own search share against
# its own consideration share over time.
q <- panel |>
  mutate(qtr = as.Date(cut(month, "quarter"))) |>
  group_by(category_name, brand_name, qtr) |>
  summarise(sos = mean(sos), soc = mean(soc), sop = mean(sop), .groups = "drop")

within_brand <- q |>
  group_by(category_name, brand_name) |>
  filter(n() >= 6, sd(sos) > 0, sd(soc) > 0) |>
  summarise(r = cor(sos, soc), .groups = "drop")

ggplot(within_brand, aes(r)) +
  geom_histogram(binwidth = .1, fill = PAL[1], color = "white") +
  geom_vline(xintercept = 0, color = "grey40") +
  labs(x = "How closely a brand's search share and consideration share move together",
       y = "Number of brands",
       title = "Brand by brand, the two barely move together",
       subtitle = paste0("Median ", round(median(within_brand$r), 2),
                         " across ", nrow(within_brand), " brands"))
```

A weak result, and it should be read as one. Some brands track well, more sit
near zero, plenty are negative.

But a correlation this size isn't a conclusion either way. Section 3 showed
quarterly survey figures still carry real noise, and two noisy series can't
correlate strongly however well connected the underlying things are. This rules
out search as a tight substitute for the survey. It doesn't say search is
uninformative.

The better move is to stop asking whether the lines move together and start
asking what the distance between them means.

# 6. Demand a brand hasn't converted yet {#esos}

Marketers already trust a version of this idea. Binet and Field's excess share
of voice rule says a brand advertising at a bigger share than its market share
tends to grow, and a brand under-investing tends to shrink. The gap between the
two points the direction. (Summary and sources in
[`assumptions.md`](assumptions.md).)

Search gives the same shape of measure without having to buy media data:

> **Excess Share of Search = share of search − share of consideration.**
> Positive means more people are actively looking for the brand than its
> consideration score implies. Negative means the brand is being claimed in
> surveys more than it is being sought out in real life.

For a CMO the two signs mean different things. A positive gap is an
opportunity: there is demand out there that the brand has not yet converted
into stated consideration, and it is usually cheaper to convert interest that
already exists than to create it. A negative gap is a warning: the brand is
living off reputation that is no longer showing up in behaviour.

One difference from Binet and Field is worth naming. Share of voice is an input
a brand controls. Share of search is an output of everything the brand has
already done. So the gap is a symptom to read, not a dial to turn.

Does today's gap tell us where consideration goes over the next six months?

```{r esos-split}
esos <- q |>
  group_by(category_name, brand_name) |>
  mutate(esos = sos - soc, d_soc = lead(soc, 2) - soc, d_sop = lead(sop, 2) - sop) |>
  ungroup() |>
  filter(!is.na(d_soc))

esos |>
  mutate(group = if_else(esos > 0, "Searched more than considered",
                         "Considered more than searched")) |>
  group_by(group) |>
  summarise(`brand-quarters` = n(),
            `change in consideration share, 6 months later` =
              scales::percent(mean(d_soc), .01), .groups = "drop")
```

```{r esos-chart, fig.height=4.6}
esos |>
  mutate(bin = cut(esos, breaks = quantile(esos, seq(0, 1, .1)),
                   include.lowest = TRUE, labels = FALSE)) |>
  group_by(bin) |>
  summarise(esos = mean(esos), d = mean(d_soc), se = sd(d_soc) / sqrt(n())) |>
  ggplot(aes(esos, d)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_errorbar(aes(ymin = d - 1.96 * se, ymax = d + 1.96 * se), width = 0,
                color = "grey75") +
  geom_point(color = PAL[1], size = 2.8) +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Excess share of search today (brands split into ten equal groups)",
       y = "Change in share of consideration six months later",
       title = "Brands searched for more than they're considered go on to gain consideration",
       subtitle = "Each point is a tenth of the sample; bars show the range the average could sit in")
```

The direction holds across the whole range. The group with the biggest search
surplus gains the most and the group with the biggest deficit loses the most.
It is not a clean straight line, and the two end groups sit furthest from it,
but they follow the same trend. That matters, because it means the result is
not being dragged along by one extreme group.

**Does the gap also predict preference?** It should, if search picks up
something real rather than a quirk of the consideration question. It does,
slightly more strongly:

```{r esos-funnel}
tibble(stage = c("Consideration", "Preference"),
       effect = c(coef(lm(d_soc ~ esos + soc, data = esos))[["esos"]],
                  coef(lm(d_sop ~ esos + sop, data = esos))[["esos"]]),
       `std errors from 0` = c(
         summary(lm(d_soc ~ esos + soc, data = esos))$coefficients["esos", 3],
         summary(lm(d_sop ~ esos + sop, data = esos))$coefficients["esos", 3])) |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

A search surplus is associated with subsequent movement at both stages —
consistent with the funnel sequence the study assumes, though association is as
far as this design goes. Nothing here rules out a campaign lifting search,
consideration and preference together on different lags.

**The obvious objection.** Share of consideration mean-reverts — a brand
sitting high has more room to fall — and the gap is *mechanically* negative
when consideration is high. So some of this could be regression to the mean in
a costume. A second version: these are shares, so they're zero-sum within a
category. One brand's gain is a rival's loss. That's what makes it competitive,
but it also means a brand can look like it's winning because a competitor
collapsed. Controlling for starting level and category:

```{r esos-models}
# Cluster-robust standard errors. Each brand contributes ~7 quarterly
# observations, and brands in a category share seasonality, competitor moves
# and fieldwork effects — so ordinary standard errors treat correlated
# observations as independent and come out too small. This is the standard CR1
# sandwich estimator, written out rather than imported so the adjustment is
# visible: bread * meat * bread, with clusters summed inside the meat.
cluster_se <- function(fit, cluster) {
  X <- model.matrix(fit); u <- residuals(fit); bread <- solve(crossprod(X))
  cluster <- as.factor(cluster)
  G <- nlevels(cluster); n <- nrow(X); k <- ncol(X)
  meat <- matrix(0, k, k)
  for (g in levels(cluster)) {
    idx <- which(cluster == g)
    score <- crossprod(X[idx, , drop = FALSE], u[idx])
    meat <- meat + tcrossprod(score)
  }
  sqrt(diag(bread %*% meat %*% bread) * (G / (G - 1)) * ((n - 1) / (n - k)))
}

esos <- esos |> mutate(brand_id = paste(category_name, brand_name))

specs <- list(
  "Gap alone"          = d_soc ~ esos,
  "+ starting level"   = d_soc ~ esos + soc,
  "+ category effects" = d_soc ~ esos + soc + factor(category_name))

# Two-way clustering (Cameron, Gelbach & Miller): brands repeat over time AND
# brands within a category move together, so I want both. The variance is
# V(brand) + V(category) - V(brand-and-category), which stops the overlap
# being counted twice.
se_two_way <- function(fit, a, b) {
  v <- cluster_se(fit, a)^2 + cluster_se(fit, b)^2 -
       cluster_se(fit, paste(a, b))^2
  sqrt(pmax(v, 0))
}

purrr::imap_dfr(specs, function(f, nm) {
  fit <- lm(f, data = esos)
  b <- coef(fit)[["esos"]]
  tibble(model = nm, effect = round(b, 3),
         `t, plain` = round(b / summary(fit)$coefficients["esos", 2], 2),
         `t, by brand` = round(b / cluster_se(fit, esos$brand_id)[["esos"]], 2),
         `t, brand + category` = round(
           b / se_two_way(fit, esos$brand_id, esos$category_name)[["esos"]], 2))
})
```

Three columns because the standard errors matter here. Each brand shows up
about seven times, and brands in the same category move together through shared
seasonality, competitor activity and the same fieldwork. Plain standard errors
treat all of those rows as independent, so they come out too small. Clustering
by brand roughly doubles them. Clustering by brand and category widens them
again. The effect holds in all three columns, but it is much less dramatic than
the plain numbers make it look.

Controlling for the starting level also shrinks the effect, so part of the raw
relationship was mean reversion.

**Where the signal actually lives.** Pushing further, the relationship is
*between* brands rather than *within* one over time:

```{r esos-between-within}
# Between: collapse each brand to its averages — one row per brand, so no
# repeated observations at all. Within: subtract each brand's own average, so
# only its quarter-to-quarter deviations remain.
between_b <- esos |> group_by(brand_id) |>
  summarise(across(c(esos, d_soc, soc), mean), .groups = "drop")
within_b <- esos |> group_by(brand_id) |>
  mutate(across(c(esos, d_soc, soc), ~ .x - mean(.x))) |> ungroup()

purrr::imap_dfr(
  list("Between brands (one row per brand)" = between_b,
       "Within a brand over time"           = within_b),
  function(d, nm) {
    fit <- lm(d_soc ~ esos + soc, data = d)
    tibble(comparison = nm, effect = round(coef(fit)[["esos"]], 3),
           t = round(summary(fit)$coefficients["esos", 3], 2))
  })
```

Collapsing to one row per brand removes the repeated-observation problem
entirely, and the effect holds: brands that sit persistently above their
category on search relative to consideration go on to gain consideration share.

Within a brand it does not hold. A single brand's quarter-to-quarter movement
around its own average gap tells you nothing useful about where that brand goes
next, and if anything it points the other way. This is disappointing, but it is
consistent with what section 5 found. A single brand's quarterly series is
mostly noise, and no amount of rearranging the data fixes that.

**So this is a positioning signal, not a timing signal.** It can tell you which
brands in a category look poised to gain or lose ground. It cannot tell you
that your gap widened this quarter and therefore your brand is about to grow.
That distinction belongs in the product spec rather than a footnote, because
getting it wrong is the easiest way to ship a metric that embarrasses
everyone.

**What the numbers mean.** The effect column is share points gained per share
point of gap. Concretely, using the between-brand figure: **a brand sitting 10
points of search share above its consideration share gains something like 0.15
of a point of consideration share over six months** — 20% heading for 20.15%.
That is a small effect and I am reporting it as one. It is a directional indicator, not a
forecast, and the gap needs to be large and persistent to mean anything, which
is exactly how excess share of voice is used in media planning.

This lines up with Tracksuit's own [Return on
Awareness](https://www.gotracksuit.com/report/return-on-awareness) work with
Google, which finds share of search running ahead of market share by 6–12
months. Mine is the same story one step earlier: search runs ahead of
consideration, which runs ahead of share.

**Does this work for small brands?** Commercially this is the question that
matters — small brands have the thinnest survey samples. My first cut said no:
splitting by awareness, the gap looked twice as strong for large brands. That
didn't survive checking. The flagged brands from §4 are small on consideration
and huge on search by construction, so they sit in the small-brand half and
drag it down.

```{r esos-by-size}
sized <- esos |>
  left_join(panel |> group_by(category_name, brand_name) |>
              summarise(awareness = mean(awareness), .groups = "drop"),
            by = c("category_name", "brand_name"))
cut_at <- median(sized$awareness, na.rm = TRUE)
eff <- function(d) round(coef(lm(d_soc ~ esos + soc, data = d))[["esos"]], 3)
small <- filter(sized, awareness < cut_at); large <- filter(sized, awareness >= cut_at)

tibble(brands = c("Smaller", "Larger"),
       `as collected` = c(eff(small), eff(large)),
       `flagged removed` = c(
         eff(anti_join(small, flagged, by = c("category_name", "brand_name"))),
         eff(anti_join(large, flagged, by = c("category_name", "brand_name")))))
```

With those seven out, small and large brands behave almost identically. The
apparent weakness on small brands was my measurement error, not a property of
small brands. Better news commercially, and a better fit with Return on
Awareness, which finds the awareness-to-search link strongest among low and
mid-awareness brands.

**The product version.** One flag on the brand dashboard — "search demand
running ahead of consideration" — plus a quadrant chart placing every brand in
the category on search surplus against consideration. The step after that is
translating the gap into dollar terms through a marketing mix model — which is where a CMO can defend or argue for an investment level in a board meeting, and where brand measurement starts behaving like financial reporting.

# 7. How long does search take to show up?

If search leads consideration, by how long? Marketing mix models answer this
with **carryover**: activity doesn't act all at once and doesn't stop dead, it
fades. I use the simplest version — carry some share of last quarter's search
into this quarter, try every share, keep whichever fits best.

```{r carryover, fig.height=4.4}
# Standard adstock:
#   effective search this quarter = (1-k) * this quarter + k * last quarter's
# k is the share carried over. The (1-k) weight matters: without it, higher k
# just inflates the numbers and "improves" the fit for arithmetic reasons.
adstock <- function(x, k) {
  out <- numeric(length(x)); out[1] <- x[1]
  for (i in seq_along(x)[-1]) out[i] <- (1 - k) * x[i] + k * out[i - 1]
  out
}

carry <- purrr::map_dfr(seq(0, 0.9, by = 0.1), function(k) {
  d <- q |>
    group_by(category_name, brand_name) |>
    filter(n() >= 6) |>
    mutate(search = as.numeric(scale(adstock(sos, k))),      # standardise within
           consideration = as.numeric(scale(soc))) |>        # brand so all count equally
    ungroup() |>
    filter(is.finite(search), is.finite(consideration))
  fit <- lm(consideration ~ search, data = d)
  tibble(carryover = k, strength = coef(fit)[["search"]], fit = summary(fit)$r.squared)
})
best <- carry |> slice_max(fit, n = 1)

ggplot(carry, aes(carryover, strength)) +
  geom_line(color = PAL[1], linewidth = .9) +
  geom_point(data = best, color = PAL[2], size = 3.6) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "How much of last quarter's search still counts this quarter",
       y = "Strength of the link to consideration",
       title = "Search feeds into consideration over roughly a quarter",
       subtitle = "Strongest when about half of last quarter's search still counts (orange point)")
```

The best fit carries about half of last quarter's search into this one, which
would put most of a search surge's association with consideration inside about
three months. Treat that as an **exploratory timing estimate**, not a
measurement: I picked the carryover rate by seeing which of ten values fitted
this same data best, so the winner is partly a property of the search rather
than of the world. It agrees with a separate, cruder check — shifting the
search series forward two to three months maximises its agreement with
consideration — which is reassuring but not independent evidence.

Two limits. The variation explained is small, so the curve tells you the
*timing* far more confidently than the *size*. And with 14 quarters per brand
this is near the edge of what the data can identify. A real marketing mix
model — media spend, saturation curves, holdout testing — is the right tool for
sizing it, and is well beyond a weekend.

**Does this contradict Return on Awareness?** No. They find search leading
*market share* by 6–12 months; I find it leading *consideration* by about
three. Consideration is earlier and faster-moving than sales, so a shorter lead
is what you'd expect. Together they describe one chain: search moves first,
consideration follows within a quarter, market share follows over the year.

# 8. Does it hold up?

The obvious way this could be an illusion: if everyone searches ice cream in
summer *and* considers ice cream in summer, every brand looks connected without
any brand-level signal.

So I broke the brand link deliberately. Each brand keeps its own consideration
history but gets a *rival's* search history from the same category — preserving
category seasonality, destroying the brand connection. A hundred times over,
that gives the range of results the data produces when there's nothing real to
find.

```{r placebo, fig.height=4.2}
# Grey band: 95% range across 100 brand-shuffled runs. Blue: the real result at
# each shift of the search series.
plc <- out("leadlag_placebo.csv")
band <- plc |> filter(run != "real") |> group_by(lag_months) |>
  summarise(lo = quantile(mean_r, .025), hi = quantile(mean_r, .975))
ggplot() +
  geom_ribbon(data = band, aes(lag_months, ymin = lo, ymax = hi), fill = "grey85") +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_line(data = filter(plc, run == "real"), aes(lag_months, mean_r),
            color = PAL[1], linewidth = 1) +
  geom_point(data = filter(plc, run == "real"), aes(lag_months, mean_r),
             color = PAL[1], size = 2.2) +
  annotate("text", x = -3, y = max(band$hi), hjust = 0, vjust = -0.5, size = 3.3,
           color = "grey45", label = "range when the brand link is broken") +
  scale_x_continuous(breaks = -3:3) +
  labs(x = "Search series shifted by k months (positive = search comes first)",
       y = "Average agreement with consideration",
       title = "The brand-level link is real; the exact lead isn't pinned down",
       subtitle = "Blue: real result. Grey: 100 runs with each brand given a rival's search history.")
```

Two claims here, with different strength.

**The brand-level signal is real.** The blue line sits above the grey band at
every shift. If this were only category seasonality, swapping in a rival's
search history would cost nothing and the blue line would sit inside the band.
It does not, so I am comfortable with this claim.

**The 2–3 month lead is not proven.** The line does peak at a two-to-three
month lead, but about one in seven shuffles produces a tilt that large. So:
direction supported, length unproven.

One limitation of the test: the shuffle stays inside the category, so if two
brands genuinely move together the placebo inherits some of that. It's a
conservative test, which is why I'm comfortable with the first claim and
cautious on the second.

The three findings do not carry equal weight, so it is worth setting them out
side by side:

| Finding | Evidence |
|---|---|
| Search carries brand-level information beyond category seasonality | Strong |
| Brands with a search surplus tend to gain consideration afterwards | Moderate — holds between brands with clustered errors, not within a brand |
| Search leads consideration by 2–3 months | Exploratory — direction consistent, size not established |

The first result is solid on this data. The second is strong enough to build a
directional product on, provided a holdout test runs alongside it. The third is
a hypothesis worth testing rather than a finding.

**Other checks.** Raising the search-volume bar doesn't change the ranking
result (0.44 to 0.47 as the cut-off rises). The gap result holds within
categories as well as across them, and at both funnel stages. Every brand whose
search identifier came back empty was automatically re-collected under its
plain name, so "no volume" means no demand rather than a lookup failure.

**Next, in priority order:** test the signal on later waves it hasn't seen;
fix the name matching in the weak categories and re-run §4; check whether the
gap predicts preference *conversion* rather than preference share.

# 9. Next potential steps

Taking this from a pilot to something usable, in roughly the order the work
would need to happen.

**A directional layer rather than a replacement.** The survey says where a
client stands. The gap says which way they are heading between waves. That is
an addition to a brand tracker rather than a cheaper version of one, and it
answers a real frustration, which is waiting a month for a number that mostly
moves on noise.

**Where it fills gaps.** It fills *time* well: continuous, so it covers the
months between waves and flags movement before the next wave confirms it. On
*brands* the answer improved once I fixed the contamination — the gap works
about as well for small brands as large ones, so this isn't a big-brands-only
tool.

The real brand-level limit is a reporting threshold, not missing demand. Trends
rounds to a 0–100 index within each comparison, so a brand competing against a
much larger rival reads as zero even with real, steady search. "Nobody searches
this brand" and "this brand is too small to show up next to Bunnings" are very
different claims, and I can't separate them with public data. Two things would:
pulling small brands against each other rather than the category leader, and
the granular volumes a Google partnership feed provides. So 1 in 5 is an upper
bound on genuine invisibility, not a measurement of it.

**On qualifying categories.** A category qualifies when most of its brands have
measurable search and search reproduces the funnel's ranking. The 0.7 rank
agreement I use as the line is an operational threshold, not a statistically
optimised one — it's roughly where a category's top few brands come out in the
right order, which is what a client actually looks at. It should be calibrated
against client outcomes rather than my judgement. About a third
pass today — and a chunk of the rest fail on name matching, not consumer
behaviour, so that share rises with better data engineering.

The case for shipping narrow and early is stronger than holding it back until
every category works. A qualification gate is friction, but it is cheap
friction, since it is a query over data Tracksuit already holds, and it buys
the freedom to put the metric in front of clients now. Shipping is also how the marketing science improves: a
client who sees their gap turn negative, acts on it, and watches the next two
waves gives me forward-looking validation I can't get from a historical
dataset. Not a clean experiment — they chose when to act — but ten of those
beat another round of my own robustness checks.

The thing worth protecting is the failure mode. Breaking things is fine when
being wrong costs a shrug. It gets expensive when a client opens the dashboard
and sees a brand called *Fruits* leading haircare, because they then stop
trusting every other number on the page. Shipping fast where the category
passes, labelling it directional, and keeping the contamination screen running
covers most of that risk.

**What search can't do.** No funnel stages. No absolute levels — a relative
index never says "23% of the category considers you". No audience cuts, at
least not from public data; Google holds demographic and location signal
internally, so a partnership feed may look very different. Those are the
survey's advantages.

**Where this fits.** Tracksuit already has a Google partnership behind Return
on Awareness. Most limits here — relative indices, sampling wobble, brands that
can't be identified — come from using free public data, not from search itself.
So the useful part isn't my specific numbers. It's the method: how to decide
where a search metric can be trusted, how to check it isn't measuring the wrong
thing, and how to test whether the effect survives the obvious objections.

**Packaging.** The directional flag belongs in the base tracker — it makes the
product stickier between waves and costs almost nothing to produce. What
justifies a higher tier is the engineering behind a trustworthy version:
curated brand-term matching for a client's own competitive set, category
qualification, alerting. Charging for the flag would be charging for something
nearly free. Charging for the work that makes it reliable is defensible — and
that work is getting cheaper, so I'd re-measure what a curated competitive set
costs to build and maintain today before pricing it off last year's number.

# Appendix A: getting the data {-}

Google Trends compares at most five terms per request and returns an index
scaled within that request only. To put every brand in a category on one scale,
the best-known brand rides along in every request as an anchor, and each batch
is rescaled to it.

The harder problem is knowing what to search for. "Koala" is a mattress company
and an animal; "Mecca" is a beauty retailer and a city. Where Google's
knowledge graph has an entry for the brand I use that instead of the words,
with hand-written rules for the rest, and every decision written to a log.

I hand-checked terms for eight categories and let automation handle the rest.
That's exactly where §4's weakest categories came from. A deliberate trade to
keep this inside a weekend, and the first thing I'd fix.

Re-running everything on raw brand names with no cleaning changes the
category-level answers very little: median rank agreement 0.55 uncurated versus
0.53 curated, across the 44 categories where both versions have enough brands.
Rank comparisons forgive a few misplaced brands. At brand level the errors are
wild in both directions — uncurated, "Ultimate" yoghurt looks thousands of
times bigger than it is, while Kärcher and Roll'd vanish over an umlaut and an
apostrophe.

So the cleaning barely matters for category research, and is essential for
anything a client sees with their own brand's name on it.

# Appendix B: method notes {-}

- **Share of search** = a brand's index ÷ the category total, monthly, among
  brands clearing the volume bar.
- **Margin of error and real movement.** Expected sampling variance per brand
  is `mean(p(1-p)/n)` across waves, where `n` is the weight of respondents who
  qualified for the category. Real movement is what's left of observed variance
  after subtracting it. Maximum correlation against a perfect predictor is the
  square root of that share. Weights are treated as respondent counts, which
  flatters the survey.
- **Carryover** uses the normalised form `(1-k)·search + k·previous`. The
  unnormalised version in most MMM material inflates the series as `k` rises,
  so a search over `k` picks whichever value manufactures the strongest trend.
  Worth knowing before trusting an off-the-shelf implementation.
- **Assumptions and theory:** [`assumptions.md`](assumptions.md).
- **Reproducing:** `Rscript run_all.R` rebuilds every number and this document
  from the committed raw data. Versions in `session-info.txt`.
